\LARGE \bf A Formal Description of Bitcoin's Script Language

### 2017-03-21 paper.tex
updated \S.II.A Approach -> Definition of Script Grammar

\subsection{Definition of Script Grammar}

\subsubsection{Syntax}
\# Bitcoin wiki -> tables describe the format\\
\# transaction is X Y, inputs are such and such\\
\# looked at enums in source code to get full list of defined opcode and their hex values\\
JOSEPH: also explain how the different syntactic groups were broken down\\

@@

We described the big-step semantics and the small-step semantics of the language using a combination of the reference implementation of the language [Ref] and of the Script page of the Bitcoin wiki [Ref].

@@

To our knowledge, no previous effort was made to provide a formal description of Script. The Bitcoin developers community relies mostly on a combination of the Bitcoin wiki Script article [2] and Script's C++ reference implementation [1] to learn the language or reason about it.

The Bitcoin wiki Script article contains an explanation of how each opcode works in plain English, and information about the standard script templates that main nodes of the network will accept. While it does provide an informal explanation for each opcode (which came in handy when writing the operational semantics), its descriptions are rather unrigourous. On a few occasions descriptions are vague and ambiguous. For example, the explanation of \texttt{OP\_LEFT} is that it takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive. Furthermore, this description cannot be used as input in automated theorem proving software.

The C++ reference implementation, by necessity, completely specifies the semantics of the language. However, the implementation also specifies details that are irrelevant to the syntactic correctness or valid/invalid judgement of a Script program. For example, the reference implementation defines a series of error codes, one for each possible cause of early termination with invalid state. This level of detail is not needed, and only complexifies reasoning about the language in the way we did in the results section. Additionally, as is the case with the existing Bitcoin wiki resource, we cannot directly use the language's implementation as axioms and rules to be used in a theorem prover.

Our proposed grammar, in constrast, unambiguously describes the behavior of Script and can (hopefully eventually) serve as basis for further analytic work.

>>>>

\subsection{Definition of Script Grammar}
The bulk of our project consists in providing a well-formed and consistent definition of the Script language. This includes a description of the syntax that the text of scripts must conform to as well as descriptions of the operational semantics by which the scripts evaluate. Due to the breadth of this component of our work, we defer exposition of the grammar to [Appendix A].

As the Bitcoin protocol lacks any formal specification by which it is defined, it is well accepted that the reference implementation defines the protocol. We therefore turned to the reference implementation to construct our description of Script's grammar---while also heavily upon the [Bitcoin Wiki] and [Mastering Bitcoin] to gain a detailed if not authoritative conceptual understanding of the relevant concepts and constructs. Insomuch as our grammar is consistent with this reference, it is correct. Insomuch as it diverges it is incorrect. Formally proving the consistency or equivalence of our grammar with the reference implementation is outside the scope of this project and is left as future work. Indeed, our grammar is purposefully incomplete as discussed below (\S II.C) due to the disproportionate expressive burden imposed by a few op-codes.
\subsubsection{Syntax}

Script is a stack-based language, bearing some semblance to Forth. It is parsed and evaluated left to right [Bitcoin Wiki/script]. The syntax of Script is defined by the binary serializations of script that reference implementation understands to be well-formed. We discover this by reviewing relevant literature (of which there is not much) and the source code itself, to glean such syntactic structures.

The "text" of a Script program is its binary serialization. In most contexts, our syntax (accurately) considers this text to be tokenized into a byte vector. Ultimately all syntactic constructs in Script are either commands which alter the state of the program in a defined manner or contextual data which parameterize the behavior of some commands.

We divide commands into "simple-word commands", termed \texttt{scom}, which are single byte commands that take no contextual program text to modify their behavior; "multiple-word commands", termed \texttt{mcom}, which rely upon contextual program text to modify their behavior---such as \texttt{OP\_IF}/\texttt{OP\_ELSE}/\texttt{OP\_ENDIF} or \texttt{OP\_PUSHDATA1}; and disabled commands, termed \texttt{dcom}, which consist of op-codes reserved for future use, disabled to remove perceived attack vectors, or reserved for testing. Commands are constructed from op-codes or words which are single byte literals with special meaning along with any relevant contextual data.

We define our syntax sufficiently to be aware of the transactions formats in which scripts are embedded in the hope that it may be used in the future to describe the semantics of commands which require such awareness---\texttt{OP\_CHECKSIG}.

@@

We provide both big-step and small-step operational semantics.

The state of the program consists entirely of the stack, $S$, the alt-stack, $AS$, and the validity $V$. The stack contains byte-vectors. The reference implementation defines the encoding used to map from byte-vectors to numbers. We take the existence of an inference rule describing said encoding for granted.

We describe the semantic meaning of each command with one or more corresponding inference rules. In order to contain the size of our description, we provide that well-formed programs with no corresponding inference rule for the execution of a command in a given state transition to the state in which the script is determined to be invalid and execution halts.

A number of op-codes have been disabled due to perceived vulnerabilities, reserved for future use, or reserved for internal use in testing. We provide inference rules for disabled op-codes, but also include them in \texttt{dcom}. We provide a contradictory inference rule for \texttt{dcom} which causes relevant commands to unequivocally evaluate to an invalid state. We provide these to allow for the modeling of the behavior of Script before these were disabled. If they are to be used, these op-codes must not be considered to be in \texttt{dcom}.

Scripts exist in transactions and op-codes exist which directly reference their transaction context. We omit these because of the disproportionate descriptive burden they introduce.

We also provide small step semantics to describe the progression of evaluation of a script program. The local and global reduction rules are compact due to the language's inherent simplicity.

@@

To our knowledge, there has been no previous effort to provide a formal specification of the Script language. As a result, the existing documentation works as a rough guideline for our work but is not complete. Explanations of how Script works are located on the Bitcoin wiki [2] and the reference implementation [1].

The Bitcoin wiki is an explanation of how each opcode works in plain English. The explanations are not rigorous, and are not in a logical formalism that allows mathematical reasoning. There are many situations where descriptions are vague and can be interpreted in more than one way.  For example, the explanation of \texttt{OP\_LEFT} takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.

The reference implementation is much more thorough that the wiki but, being C++ code, it is much harder to interpret and is not in a format that allows for logical reasoning. If we turn to Forth to find something related, we can find standard specifications like the Forth-83 standard (http://forth.sourceforge.net/standard/fst83/fst83-12.htm). However even these specifications are given in plain English.
